Baseline centering problem

### 2018-12-10 paper/tangotd.tex

\begin{defi}
A matrix $A$ is \emph{stable} if one of the following equivalent
conditions are satisfied:
\begin{itemize}
\item All the eigenvalues of $A$ have positive real part.
\item The matrix $(\Id-\eta A)$ has spectral radius $<1$ for small enough
$\eta$.
\item For $\eta$ small enough, iterating $\theta\gets (\Id-\eta A)\theta$
converges to $0$.
\item The differential equation $\theta'=-A\theta$ converges to $0$ for
any initial value.
\item There exists a symmetric, positive definite matrix $L$ (Lyapunov
function) such that $\transp{\theta}L\theta$ is decreasing along the
solutions of the differential equation $\theta'=-A\theta$.
\item There exists a symmetric, positive definite matrix $L$ such that
\begin{equation}
\transp{\theta} L A \theta>0
\end{equation}
for all $\theta\neq 0$. (This is the same $L$ as in the previous
condition.)
\end{itemize}
\end{defi}

@@

\begin{table}
\begin{tabular}{cccc}
Algorithm & Conditions & Preconditioner & Lyapunov
\\
\hline
\emph{Linear $V$} \\
TD	& On-policy & $B=\Id$ & $L=\Id$ \\
TD & Tabular & $B=\Id$ & $L=\mu/\rho$ \\
GTD & 2ts & $B=\transp{A}$ & $L=\Id$\\
– & On-policy, 2ts & $B=(\transp{\Phi}\rho\Phi)^{-1}$ & $L=B^{-1}$ \\
– & Overparameterized, 2ts & $B=(\transp{\Phi}\rho\Phi)^{-1}$ &
$L=\transp{\Phi}\mu\Phi$ \\
TDC, GTD2 & 2ts & $B=\transp{A}(\transp{\Phi}\rho\Phi)^{-1}$ &
$L=\Id$ \\
2nd-order & On-policy, 2ts & $B=A^{-1}$ & $L=\Id$\\
2nd-order & 2ts & $B=\transp{A}(A\transp{A})^{-1}$ & $L=\Id$\\
– & 2ts & $B=\transp{A}(A+\transp{A})^{-1}$ & $L=\Id$\\
\hline
\emph{Non-linear $V$} \\
TD & On-policy, reversible & any $B\succ 0$ & $L=\mu(\mu(\Id-P))^{-1}\mu$
\\
– & On-policy, overparameterized, 2ts & $B=(\transp{\Phi}S\Phi)^{-1}$ &
$L= \mu S^{-1} \mu$ \\
Non-lin GTD2 & 2ts, different objective &
$B=\transp{A}(\transp{\Phi}\rho\Phi)^{-1}$  & $L=\Id$
\\\hline
\end{tabular}
\caption{Various cases of convergence of TD-like algorithms, the
preconditioner $B$ for the TD step, and the matrix $L$ that proves
convergence via the Lyapunov criterion. 2ts=two-timescale.}
\end{table}

\paragraph{Making the TD iteration stable.} In linear TD the matrix $A$ is
\begin{equation}
A\deq \transp{\Phi}\rho (\Id-P)\Phi
\end{equation}
and we want to solve
\begin{equation}
A\theta=b\,\qquad b\deq \transp{\Phi}\rho R
\end{equation}
If $A$ is stable, then iterating $\theta\gets \theta -\eta(A\theta-b)$ will
converge.

We know $A$ is stable in several cases: (TODO here I just ignore constant
functions)
\begin{enumerate}
\item On-policy: $\rho=\mu$, the invariant distribution. Then we know that $\rho(\Id-P)$ is positive
(Dirichlet form: $\transp{f}\rho(\Id-P)f=\E(f(s_t)-f(s_{t+1}))^2$). This implies that $A=\transp{\Phi}\rho (\Id-P)\Phi$ is
positive too. So we can take $L=\Id$ and we recover convergence of linear
TD on-policy.
\item In the tabular case: $\Phi=\Id$. Indeed we have $A=\rho (\Id-P)$,
which is not positive in general. But $\mu (\Id-P)$ is positive, so we can take
$L=\mu/\rho$ so that $LA$ is positive. Thus, tabular TD works
off-policy.
\end{enumerate}

>>>>

\begin{defi}[ (Stable matrix)]
A matrix $A$ is \emph{stable} if one of the following equivalent
conditions are satisfied:
\begin{itemize}
\item All the eigenvalues of $A$ have positive real part.
\item The matrix $(\Id-\eta A)$ has spectral radius $<1$ for small enough
$\eta$.
\item For $\eta$ small enough, iterating $\theta\gets (\Id-\eta A)\theta$
converges to $0$.
\item The differential equation $\theta'=-A\theta$ converges to $0$ for
any initial value.
\item There exists a symmetric, positive definite matrix $L$ (Lyapunov
function) such that $\transp{\theta}L\theta$ is decreasing along the
solutions of the differential equation $\theta'=-A\theta$.
\item There exists a symmetric, positive definite matrix $L$ such that
\begin{equation}
\transp{\theta} L A \theta>0
\end{equation}
for all $\theta\neq 0$. (This is the same $L$ as in the previous
condition.)
\end{itemize}
\end{defi}

@@

\begin{table}
\begin{tabular}{cccc}
Algorithm & Conditions & Preconditioner & Lyapunov
\\
\hline
\emph{Linear $V$} \\
TD	& On-policy & $B=\Id$ & $L=\Id$ \\
TD & Tabular & $B=\Id$ & $L=\mu/\rho$ \\
GTD & 2ts & $B=\transp{A}$ & $L=\Id$\\
– & On-policy, 2ts & $B=(\transp{\Phi}\rho\Phi)^{-1}$ & $L=B^{-1}$ \\
– & Overparameterized, 2ts & $B=(\transp{\Phi}\rho\Phi)^{-1}$ &
$L=\transp{\Phi}\mu\Phi$ \\
TDC, GTD2 & 2ts & $B=\transp{A}(\transp{\Phi}\rho\Phi)^{-1}$ &
$L=\Id$ \\
2nd-order & On-policy, 2ts & $B=A^{-1}$ & $L=\Id$\\
2nd-order & 2ts & $B=\transp{A}(A\transp{A})^{-1}$ & $L=\Id$\\
– & 2ts & $B=\transp{A}(A+\transp{A})^{-1}$ & $L=\Id$\\
\hline
\emph{Non-linear $V$} \\
TD & On-policy, reversible & any $B\succ 0$ & $L=\mu(\mu(\Id-P))^{-1}\mu$
\\
– & On-policy, overparameterized, 2ts & $B=(\transp{\Phi}S\Phi)^{-1}$ &
$L= \mu S^{-1} \mu$ \\
Non-lin GTD2 & 2ts, different objective &
$B=\transp{A}(\transp{\Phi}\rho\Phi)^{-1}$  & $L=\Id$
\\\hline
\end{tabular}
\label{tab:TDlist}
\caption{Various cases of convergence of TD-like algorithms, the
preconditioner $B$ for the TD step, and the matrix $L$ that proves
convergence via the Lyapunov criterion. 2ts=two-timescale.}
\end{table}

\paragraph{Making the TD iteration stable.} In linear TD the matrix $A$ is
\begin{equation}
A\deq \transp{\Phi}\rho (\Id-P)\Phi
\end{equation}
and we want to solve
\begin{equation}
A\theta=b\,\qquad b\deq \transp{\Phi}\rho R
\end{equation}
If $A$ is stable, then iterating $\theta\gets \theta -\eta(A\theta-b)$ will
converge. (TODO: I will ignore the problem with constant functions for
$\gamma=1$.)

We now review several cases in which $A$ is known to be stable, or cases
in which a modified TD algorithm makes the iteration stable (by
preconditioning the gradient descent by a matrix $B$). See also
Table~\ref{tab:TDlist}.

\begin{enumerate}
\item On-policy: $\rho=\mu$, the invariant distribution. Then we know that $\rho(\Id-P)$ is positive
(Dirichlet form: $\transp{f}\rho(\Id-P)f=\E(f(s_t)-f(s_{t+1}))^2$). This implies that $A=\transp{\Phi}\rho (\Id-P)\Phi$ is
positive too. So we can take $L=\Id$ and we recover convergence of linear
TD on-policy.
\item In the tabular case: $\Phi=\Id$. Indeed we have $A=\rho (\Id-P)$,
which is not positive in general. But $\mu (\Id-P)$ is positive, so we can take
$L=\mu/\rho$ so that $LA$ is positive. Thus, tabular TD works
off-policy.
\end{enumerate}
